---
title: "Homework 3 Solutions"
author: "Drew Crossett"
date: "9/22/2023"
output: word_document
---

## Introduction

```{r}
x <- 1:5
y <- seq(from = 2, to = 10, by = 2)
z <- c(1, 2, 4, 7, 11)
names.1 <- c("Angela", "Brie", "Drew", "Annie", "Ally", "Matt")
names.2 <- c("Andrew", "Krista", "Jesse", "Paul", "Adam", "Monica")
```

## (a)

The following code adds 2 to each element of vector ```z```

```{r}
z + 2
```

## (b)

The following code multiplies the values of ```x``` and ```y``` element-by-element. Recall that R uses the recycling rule if they are not the same length!

```{r}
x*y
```

## (c)

This code gives me the first, second, and fifth elements from the vector ```names.1```

```{r}
names.1[ c(1, 2, 5) ]
```

## (d)

The following code checks every element of ```y``` to see if it's equal to 4. If it is then it will print ```TRUE``` (and ```FALSE``` otherwise).

```{r}
y == 4
```

## (e)

The following code combines the elements from vectors ```x``` and ```y``` into a single vector. The order matters!

```{r}
c(x, y)
```

## (f)

The following code repeats the second and third values of ```x``` three times.

```{r}
rep(x = x[ 2:3 ], times = 3)
```

## (g)

The following code checks every element of ```z``` to see if it's less than or equal to 7. If it is then it prints ```TRUE``` (```FALSE``` otherwise)

```{r}
z < 7
```

## (h)

Print the values of ```y``` such that those elements were ```TRUE``` from part (g).

```{r}
y[ z < 7 ]
```

## (i)

The following code compares everything in  ```names.1``` to the elements in ```names.2```element-by-element. If the value in ```names.1``` comes before the same element in ```names.2``` alphabetically then it prints ```TRUE``` (and ```FALSE``` otherwise).

```{r}
names.1 < names.2
```

## (j)

The following code raises each element of ```z``` to the third power.

```{r}
z^3
```