---
title: "STA552 Assignment 2 - Imputation"
author: " Junjie Mei"
date: " 02/25/2025 "
output:
  html_document: 
    toc: yes
    toc_depth: 4
    toc_float: yes
    number_sections: yes
    toc_collapsed: yes
    code_folding: hide
    code_download: yes
    smooth_scroll: yes
    theme: lumen
  word_document: 
    toc: yes
    toc_depth: 4
    fig_caption: yes
    keep_md: yes
  pdf_document: 
    toc: yes
    toc_depth: 4
    fig_caption: yes
    number_sections: no
    fig_width: 3
    fig_height: 3
editor_options: 
  chunk_output_type: inline
---

```{=html}

<style type="text/css">

/* Cascading Style Sheets (CSS) is a stylesheet language used to describe the presentation of a document written in HTML or XML. it is a simple mechanism for adding style (e.g., fonts, colors, spacing) to Web documents. */

h1.title {  /* Title - font specifications of the report title */
  font-size: 22px;
  font-weight: bold;
  color: DarkRed;
  text-align: center;
  font-family: "Gill Sans", sans-serif;
}
h4.author { /* Header 4 - font specifications for authors  */
  font-size: 18px;
  font-weight: bold;
  font-family: system-ui;
  color: navy;
  text-align: center;
}
h4.date { /* Header 4 - font specifications for the date  */
  font-size: 18px;
  font-family: system-ui;
  color: DarkBlue;
  text-align: center;
  font-weight: bold;
}
h1 { /* Header 1 - font specifications for level 1 section title  */
    font-size: 22px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: center;
    font-weight: bold;
}
h2 { /* Header 2 - font specifications for level 2 section title */
    font-size: 20px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
    font-weight: bold;
}

h3 { /* Header 3 - font specifications of level 3 section title  */
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}

h4 { /* Header 4 - font specifications of level 4 section title  */
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: left;
}

body { background-color:white; }

.highlightme { background-color:yellow; }

p { background-color:white; }

</style>
```


```{r setup, include=FALSE}
# code chunk specifies whether the R code, warnings, and output 
# will be included in the output files.
if (!require("knitr")) {
   install.packages("knitr")
   library(knitr)
}
if (!require("tidyverse")) {
   install.packages("tidyverse")
library(tidyverse)
}
if (!require("ggplot2")) {
   install.packages("ggplot2")
   library(knitr)
}
if (!require("dplyr")) {
   install.packages("dplyr")
library(dplyr)
}
if (!require("VIM")) {
   install.packages("VIM")
   library(VIM)
}
if (!require("mice")) {
   install.packages("mice")
   library(mice)
}

## 
knitr::opts_chunk$set(echo = TRUE,   # include code chunk in the output file
                      warning = FALSE,# sometimes, you code may produce warning messages,
                                      # you can choose to include the warning messages in
                                      # the output file. 
                      results = TRUE, # you can also decide whether to include the output
                                      # in the output file.
                      message = FALSE,
                      comment = NA
                      )  
```


## Introduction

This dataset is collected internally from a bank about  the monthly loan payment amount and the current bank accounts' balances from the bank customers with either a car loan or a personal loan. The total sample size is 1000. There are 16 variables, including 13 numerical variables and 3 categorical variable:

 * Checking_amount: current Checking account's balance
 * Saving_amount: current saving account's balance
 * Amount: current loan's monthly payment
 * Term: Current loan's payment times
 * Credit_score: the customer's credit score
 * Gender: Customer's gender (Male and Female)
 * Marital_status: Customer's marital status (Single and Married)
 * Emp_status: Customer's employed status (employed and unemployed)
 * Car_loan: whether this customer has a car loan in this bank (1 and 0)
 * Personal_loan: whether this customer has a personal loan in this bank (1 and 0)
 * Home_loan: whether this customer has a home loan in this bank (1 and 0)
 * Education_loan: whether this customer has a education loan in this bank (1 and 0)
 * Age: this customer's age
 * No_of_credit_acc: how many credit card account does this customer have in this bank 

We created 5 featured variables:

 * TotalAmount: Numerical variable, it's the combination of Saving_Amount and Checking-Amount.
 * ActualAmount: Numerical variable, it's the combination of Saving_Amount and Checking-Amount minus Monthly payment Amount.
 * Gender_Marital: Categorical variable, it's the combination of the clients' Gender and marital status.
 * Gender_Dummy: convert the categorical variable Gender into a dummy variable (1: Male and 0: Female)
 * Marital_status_Dummy: convert the categorical variable Marital_status into a dummy variable (1: Married and 0: Single)

 
The purpose of this assignment is to use different impuation approach for preparation for the data sets ready for machine learning. This dataset has 1000 rows with no missing values. So we randomly replace 15% of values of the following variables with missing values (NA): TotalAmount, ActualAmount, Amount, Gender_Marital, Gender, Marital_status. 


```{r}
# load CSV bank data
# setwd("C:/Users/Junjie Mei/Desktop/WCU/2025Spring/STA552/Week2")
mpg <- read.csv("https://jm1006958.github.io/STA552/BankLoanDefaultDataset.csv")

#generate  feature  variables 
mpg <- mutate(mpg, TotalAmount = Saving_amount + Checking_amount )
mpg <- mutate(mpg, ActualAmount = Saving_amount + Checking_amount - Amount ) 
mpg <- mutate(mpg, Gender_Marital = paste(Gender, Marital_status, sep = " "))

mpg$Gender_Dummy <- ifelse(mpg$Gender == "Male", 1, 0)
mpg$Marital_status_Dummy <- ifelse(mpg$Marital_status == "Married", 1, 0)
mpg$Emp_status_Dummy <- ifelse(mpg$Emp_status == "employed", 1, 0)
#print(mpg)

# add 150 missing values to 3 numerical variables and 3 categorical variables
#library(tidyverse)
TotalAmount.missing.id <- sample(1:1000, 150, replace = FALSE)
ActualAmount.missing.id <- sample(1:1000, 150, replace = FALSE) 
Amount.missing.id <- sample(1:1000, 150, replace = FALSE)
Gender_Marital.missing.id <- sample(1:1000, 150, replace = FALSE)
Gender.missing.id <- sample(1:1000, 150, replace = FALSE)
Marital_status.missing.id <- sample(1:1000, 150, replace = FALSE)
mpg$TotalAmount[TotalAmount.missing.id] <- NA
mpg$ActualAmount[ActualAmount.missing.id] <- NA
mpg$Amount[Amount.missing.id] <- NA
mpg$Gender_Marital[Gender_Marital.missing.id] <- NA
mpg$Gender[Gender.missing.id] <- NA
mpg$Marital_status[Marital_status.missing.id] <- NA

ggplot(mpg, aes(x = Amount)) +
  geom_histogram(aes(y = ..density..), bins = 20, fill = "steelblue", color = "black", alpha = 0.7) +
  geom_density(color = "red", size = 1) +  # Overlay density plot
  ggtitle("Distribution of Amount") +
  xlab("Amount") +
  ylab("Density") +
  theme_minimal()

ggplot(mpg, aes(x = TotalAmount)) +
  geom_histogram(aes(y = ..density..), bins = 20, fill = "steelblue", color = "black", alpha = 0.7) +
  geom_density(color = "red", size = 1) +  # Overlay density plot
  ggtitle("Distribution of TotalAmount") +
  xlab("TotalAmount") +
  ylab("Density") +
  theme_minimal()

ggplot(mpg, aes(x = ActualAmount)) +
  geom_histogram(aes(y = ..density..), bins = 20, fill = "steelblue", color = "black", alpha = 0.7) +
  geom_density(color = "red", size = 1) +  # Overlay density plot
  ggtitle("Distribution of ActualAmount") +
  xlab("ActualAmount") +
  ylab("Density") +
  theme_minimal()
#print(mpg)
pairs(mpg[, c("ActualAmount", "Gender_Dummy", "Marital_status_Dummy", "TotalAmount", "Amount")], main = "Pairwise Plot of Selected Variables")

ggplot(mpg, aes(x = Gender)) +
  geom_bar(fill = "steelblue", color = "black", alpha = 0.7) +
  ggtitle("Distribution of Gender") +
  xlab("Gender") +
  ylab("Count") +
  theme_minimal()

ggplot(mpg, aes(x = Marital_status)) +
  geom_bar(fill = "steelblue", color = "black", alpha = 0.7) +
  ggtitle("Distribution of Marital_status") +
  xlab("Marital_status") +
  ylab("Count") +
  theme_minimal()

ggplot(mpg, aes(x = Gender_Marital)) +
  geom_bar(fill = "steelblue", color = "black", alpha = 0.7) +
  ggtitle("Distribution of Gender_Marital") +
  xlab("Gender_Marital") +
  ylab("Count") +
  theme_minimal()
```


## Replacement Imputation for Categorical Features

Use the k-nearest neighbors (KNN) algorithm to impute missing values of the "Gender_Marital" variable by assigning the value most common among the k-nearest neighbors surrounding the point with the missing data

```{r}

ggplot(mpg, aes(x = Gender_Marital)) +
  geom_bar(fill = "steelblue", color = "black", alpha = 0.7) +
  ggtitle("Distribution of Gender_Marital Before KNN Imputation") +
  xlab("Gender_Marital") +
  ylab("Count") +
  theme_minimal()

mpg <- kNN(mpg, variable = "Gender_Marital", k = 4)

ggplot(mpg, aes(x = Gender_Marital)) +
  geom_bar(fill = "steelblue", color = "black", alpha = 0.7) +
  ggtitle("Distribution of Gender_Marital after KNN Imputation") +
  xlab("Gender_Marital") +
  ylab("Count") +
  theme_minimal()
```

There is no female married category, indicating in a married family, only husband will work on car loan and personal loan with this bank on this bank-covered area.


## Regression-based Imputation for Numerical Features
Create a subset containing the feature variables associated with the target continuous response variable. Fit a linear regression model using the complete observations from this subset. For instances where the response variable has missing values, predict the missing values using the model’s predicted output, adding random error to account for variability. Note that the feature variables can be a mix of categorical and numerical data types.
```{r}
	
pred.Am = lm(Amount ~ ActualAmount, data = mpg)
newdata = mpg[is.na(mpg$Amount),]
pred.Amount = predict(pred.Am, newdata = newdata)
m0 = sum(is.na(mpg$Amount))  
pred.resid = resid(pred.Am)  
pred.yrand = pred.Amount + sample(pred.resid, m0, replace = TRUE)

# Ensure no missing values in x-axis data
complete_cases <- !is.na(mpg$ActualAmount) & !is.na(mpg$Amount)
plot(mpg$ActualAmount[complete_cases], mpg$Amount[complete_cases], main = "ActualAmount vs Amount")

abline(pred.Am, col = "steelblue", lty = 2, lwd = 2)

# Plot only for missing values
if (m0 > 0) {
  points(newdata$ActualAmount, pred.Amount, pch=19, col = "red")
  points(newdata$ActualAmount, pred.yrand, pch=19, col = "blue")
}

legend("topleft", c("regression imputation", "random regression imputation"),
       col=c("red", "blue"), pch=rep(19,2), bty="n", cex = 0.8)
```

There is no significant difference on on the account balances and the residual balances after monthly payments between male customers and female customers.


## Multiple Imputation
Develop a regression model, either linear or nonlinear, as appropriate.
Present the results of the regression analysis using the MICE imputation procedure.
Apply the same regression model to the imputed dataset obtained in Sections 2 and 3, and compare its performance with the results from the MICE procedure.
Provide practical insights and recommendations based on the analysis.

```{r}
dat <- mpg[,c("TotalAmount","ActualAmount")]  # Example data set with missing values
imp5 <- mice(dat, method = c("pmm","pmm"), m = 5, maxit = 10, seed = 123, print=F)
plot(imp5)

model5 <- with(imp5, lm(TotalAmount ~ ActualAmount))  # The statistical model to assess
# the relationship between *Ozone*
# and *Wind* and *Temp*.
summary.stats = summary(model5)               # display the regression results of 
# individual imputed data set.
summary.stats 
summary(pool(model5))
```

TotalAmount and ActualAmount are positively correlated based on the paired plot above. Due to my problem, I did not have time to finish the 4 tasks for this "Multiple Imputation" question, but I will try my best to complete them later. Thanks!



